Certified incremental program transformations

### 2012-05-01 doc/papers/LFMTP2012/paper.tex
paper

@@

As is customary, we adopt the Barendregt convention for naming bound
variables, write $\meta X$ for $\smeta X \msubstnil$, $A\to B$ for
$\prd x A B$ when $x\notin \FV{B}$ and adopt list notations for $S,
\sigma, \Gamma, \Sigma$ when convenient.

@@

A \emph{signature} declares object and type constants to be used in
objects:
$$
\Sigma \gequal
\gnil \gor
\gcons \Sigma {\cst c : A} \gor
\gcons \Sigma {\cst a : K} \qquad
\text{Signature}
$$

The algorithm is presented in Fig. \ref{fig:typing-algo} as a
syntax-directed inference system. All judgements are of the form $\gj
U V W$ where $U$ and $V$ are inputs, $V$ is the induction hypothesis
and $W$ the output. They are parameterized by a constant and implicit
signature $\Sigma$.

@@

@@

\subsection{The computation calculus}

We now turn to the definition of our computation calculus on LF
objects. This language is untyped. % TODO??
\begin{align*}
  T &\gequal
  F \gor
  \matchin T \Gamma \casenobar P T\ |\ \ldots\ \case P T \\
  P &\gequal \app {\cst c} {\vec x}
\end{align*}

We extend our notion of signature to allow object constants to embed a
computational behaviour.

$$  \Sigma \gequal
  \ldots \gor
  \gcons \Sigma {\cst c : A = T}
  \qquad
  \text{Signature}
$$

>>>>

\subsection{The LF representation language}
\subsection{Storing computation results}
\subsection{Naming intermediate results}

@@

As is customary, we adopt the Barendregt convention for naming bound
variables, write $A\to B$ for $\prd x A B$ when $x\notin \FV{B}$,
$\meta X$ for $\smeta X \msubstnil$, and adopt indifferently list or
map notations for $\sigma$ and $\Gamma$, \ie\ write $\sigma(x) = M$ or
$\Gamma(x) = A$.

@@

A \emph{signature} declares object and type constants to be used in
objects. To each object constant we assign an annotation $L$ setting
the slicing behaviour of the constant during type-checking: either
sliceable $\annsliceable$ (each application of this constant will have
its own name in the repository), non-sliceable $\annnonsliceable$ (we
don't create a name for applications of this constant), or defined
$\anndefined{F}$ (see \ref{sec:computational-calculus}).
\begin{align*}
  L &\gequal
  \annsliceable \gor
  \annnonsliceable \gor
  \anndefined{F}
  \\
  \Sigma &\gequal
  \gnil \gor
  \gcons \Sigma {\cst c :^L A} \gor
  \gcons \Sigma {\cst a :^L K}
  &\text{Signature}
\end{align*}
We write $\Sigma(\cst c) = A; L$ for $(\cst c :^L A) \in \Sigma$
suggesting the implementation of signatures as maps.

The algorithm is presented in Fig. \ref{fig:typing-algo} as a
syntax-directed inference system. All judgements are of the form $\gjo
U V W$ where $U$ and $V$ are inputs, $V$ is the induction hypothesis
and $W$ the output. They are parameterized by a constant and implicit
signature $\Sigma$.

@@

  \fbox{$\jf\mr\Gamma F \mmr M A$}
  \qquad
  {Application typing}

    % Non-sliceable
    \infer{
      \jh\mr\Gamma H A \annnonsliceable
      \and
      \jl\mr\Gamma A S \mmr {S'} {A'}
    }{
      \jf\mr\Gamma {\app H S} \mmr {\app H {S'}} {A'}
    }

    % Sliceable
    \infer{
      \jh\mr\Gamma H A \annsliceable
      \and
      \jl\mr\Gamma A S \mmr {S'} {A'}
      \and
      \jp\mmr\Gamma {\app H S} A \mmmr
    }{
      \jf\mr\Gamma {\app H S} \mmmr {\hat\mmmr} {A'}
    }

  \fbox{$\jh\mr\Gamma H A L$}
  \qquad
  {Head typing}

  \begin{mathpar}
    % constant
    \infer{
      \Sigma(\cst c) = A; L
    }{
      \jh\mr\Gamma {\cst c} A L
    }

    % variable
    \infer{
      \Gamma(\var x) = A
    }{
      \jh\mr\Gamma {\var x} A \annnonsliceable
    }
  \end{mathpar}

  \fbox{$\jp\mr\Gamma M A \mmr$}
  \qquad
  {Push}

  \begin{mathpar}
    \infer{
      \jstr{\Gamma\vdash\tp M A}{\Gamma'\vdash\tp {M'} {A'}}\sigma
      \and
      \text{$\meta X$ fresh in $\Delta$}
    }{
      \jp{(\Delta; \smeta Z \sigma)}\Gamma M A {(\mapadd\Delta{\meta
          X}{(\Gamma'\vdash M':A')}; \smeta X \sigma)}
    }
  \end{mathpar}

@@

\subsection{The computation calculus}
\label{sec:computational-calculus}

We now turn to the definition of our computation calculus on LF
objects. This language is untyped. % TODO??
\begin{align*}
  T &\gequal
  F \gor
  \matchin T \Gamma \casenobar {\app {\cst c} {\vec x}} T\ |\ \ldots\
  \case {\app {\cst c} {\vec x}} T
\end{align*}
